A brief review of Quantum Mechanics for Computer Scientists

### 2023-06-18 documentation/article02/text/article.tex
Ket notation fix

			\begin{equation}
				|K\rangle = \vec{K} = \begin{bmatrix}
					a \\
					b \\
					c \\
					d
				\end{bmatrix}
				\label{eq:ketNotationVector}
			\end{equation}

@@

			\begin{equation}
				| \psi \rangle = \alpha . | V_1 \rangle + \beta . | V_2 \rangle
				\label{eq:ketNotationGeneric}
			\end{equation}

@@

			\par For instance, consider a three-level system. We can define a basis as ${|V_0\rangle, |V_1\rangle, |V_2\rangle}$, where $|V_0\rangle$, $|V_1\rangle$, and $|V_2\rangle$ represent the vectors associated with the respective levels.

@@

			\begin{equation}
				|\psi\rangle = \alpha |V_0\rangle + \beta |V_1\rangle + \gamma |V_2\rangle \qquad,
				\label{eq:multipleLevelState}
			\end{equation}

@@

			\par Being $N=$ the number of levels and $a \in (\alpha, \beta, \gamma, ... , \omega)$ sometimes the equation \ref{eq:multipleLevelState} can be represented as a sum:

			\begin{equation}
				|\psi\rangle = \sum_{i=1}^{N} a_i . |V_i\rangle = 
				\begin{bmatrix}
					a_{1} \\
					a_{2} \\
					\vdots \\
					a_{n}
				\end{bmatrix}
				\label{eq:multipleLevelStateSum}
			\end{equation}

			\par Some authors prefers to use $|0\rangle, |1\rangle, |2\rangle, \dots , |n\rangle / n < N$ instead of $|V_0\rangle,|V_1\rangle, \dots, |V_n\rangle / n < N$.\newline

@@

			\begin{equation}
				\label{eq:linearCombination}
				\begin{aligned}
					|\psi\rangle =
					\begin{bmatrix}
						a_{1} \\
						a_{2} \\
						\vdots \\
						a_{n}
					\end{bmatrix} =
					a_1 . \begin{bmatrix}
						1\\
						0\\
						\vdots \\
						0
					\end{bmatrix}+
					a_2 . \begin{bmatrix}
						0\\
						1\\
						\vdots \\
						0
					\end{bmatrix}+
					\dots + a_n . \begin{bmatrix}
						0\\
						0\\
						\vdots \\
						1
					\end{bmatrix}
				\end{aligned}
			\end{equation}

@@

			\par Briefly, given a ket $|\psi\rangle$ (as shown in equation \ref{eq:ket}), its corresponding bra is denoted by the conjugate transpose of $|\psi\rangle$ (equation \ref{eq:bra}).

			\begin{equation}
				\label{eq:ket}
				|\psi\rangle = 
				\begin{bmatrix}
					u + vo \\
					k + ji \\
					o + li \\
					d + mi
				\end{bmatrix} \in \mathbb{C}
			\end{equation}

			\begin{equation}
				\label{eq:bra}
				\langle\psi| = [u - vo, k - ji, o - li, d - mi]
			\end{equation}

@@

			\par As we already know, a vector of all probability amplitudes, denoted as $|\psi\rangle$, is given by equation \ref{eq:multipleLevelStateKet}. So, for these "ket" states, there is an associated "bra" notation defined in equation \ref{eq:multipleLevelStateBra} \cite{notacaoDirac}

			\begin{equation}
				|\psi\rangle = \sum_{i=1}^{N} a_i . |V_i\rangle = 
				\begin{bmatrix}
					a_{1} \\
					a_{2} \\
					\vdots \\
					a_{n}
				\end{bmatrix}
				\label{eq:multipleLevelStateKet}
			\end{equation}

			\begin{equation}
				\langle\psi| = \sum_{j=1}^{N} a_j^* . \langle V_j| = [a_1^*, a_2^*, \dots, a_n^*]
				\label{eq:multipleLevelStateBra}
			\end{equation}

@@

			\begin{equation}
				\begin{aligned}
					\langle\psi|\psi\rangle = [a_1^*, a_2^*, \dots, a_n^*] . \begin{bmatrix}
						a_{1} \\
						a_{2} \\
						\vdots \\
						a_{n}
					\end{bmatrix} = &\\ 
					a_1^*.a_1 + a_2^*.a_2 + \dots + a_n^*.a_n = &\\
					P(S_1) + P(S_2) + \dots + P(S_n) = &\\
					\sum_{i=1}^{N} P(S_i)
				\end{aligned}
				\label{eq:multipleLevelStateBraKet}
			\end{equation}

			\par Bra notation allows for the calculation of inner products between vectors. The inner product of a bra vector $\langle \psi |$ and a ket vector $| \phi \rangle$ is denoted as $\langle \psi | \phi \rangle$. The inner product yields a complex number, providing information about the similarity or overlap between the two vectors.\newline

			\par For instance, given a vector 
				$|\alpha\rangle= 
				\begin{bmatrix}
					1 \\
					0
				\end{bmatrix}
				$
				 and 
				 $|\beta\rangle = \begin{bmatrix}
					0\\
					1
				\end{bmatrix}$ 
				using the "BraKet" operators to represent the inner product, it will end up like in equation \ref{eq:braket}.

			\begin{equation}
				\label{eq:braket}
				\langle \alpha | \beta \rangle = 
				[1, 0] .
				\begin{bmatrix}
					0 \\
					1
				\end{bmatrix} = 1.0 + 0.1 = 0
			\end{equation}

@@

		\begin{equation}
			| \psi \rangle = \alpha . | 0 \rangle + \beta . | 1 \rangle = 
			\alpha . \begin{bmatrix}
				1 \\
				0
			\end{bmatrix} + \beta .	\begin{bmatrix}
				0 \\
				1
			\end{bmatrix} = 
			\begin{bmatrix}
				\alpha \\
				\beta
			\end{bmatrix}
			\label{eq:qubit}
		\end{equation}

		\par The coefficients $\alpha$ and $\beta$ are complex numbers such that ${|\alpha|}^2 + {|\beta|}^2 = 1$, the vector 
		$\begin{bmatrix}
			\alpha \\
			\beta 
		\end{bmatrix}$ is the qubit itself and, by convention, $| 1 \rangle = 
		\begin{bmatrix}
			0 \\
			1
		\end{bmatrix} = 1$ and $| 0 \rangle = 
		\begin{bmatrix}
			1 \\
			0
		\end{bmatrix} = 0$

@@

		\par The Equation \ref{eq:qubit} can be interpreted as follows: The \textbf{non-measured} state of $\psi$ within the vector space defined by vectors $| 1 \rangle$ and $| 0 \rangle$ is a linear combination of them  ($\alpha$ times vector $| 0 \rangle$ plus $\beta$ times vector $| 1 \rangle$).\newline

@@

			\par As mentioned earlier, quantum gates are represented as matrices, and their number is infinite because the values of the matrix cells can vary across the real numbers ($\mathbb{R}$). Although there is a wide variety of quantum gates and considering $| \psi \rangle = \alpha . | 0 \rangle + \beta . | 1 \rangle$, here are a few examples.\newline

@@

			\begin{equation}
				\label{eq:indentityApplyed}
				I|\psi\rangle = \begin{bmatrix}
					1& 0 \\
					0& 1
				\end{bmatrix} . \begin{bmatrix}
					\alpha \\
					\beta
				\end{bmatrix} = \begin{bmatrix}
					\alpha \\
					\beta
				\end{bmatrix} = |\psi\rangle
			\end{equation}

@@

			\begin{equation}
				\label{eq:flipApplyed}
				X|\psi\rangle = \begin{bmatrix}
					0& 1 \\
					1& 0
				\end{bmatrix} . \begin{bmatrix}
					\alpha \\
					\beta
				\end{bmatrix} = \begin{bmatrix}
					\beta \\
					\alpha
				\end{bmatrix}
			\end{equation}

@@

			\begin{equation}
				\label{eq:hadamardApplyed}
				\begin{aligned}
					&H|\psi\rangle = \dfrac{1}{\sqrt{2}}.\begin{bmatrix}
						1& 1 \\
						1& -1
					\end{bmatrix} . \begin{bmatrix}
						\alpha \\
						\beta
					\end{bmatrix} = \\
					&\begin{bmatrix}
						\dfrac{1}{\sqrt{2}} &\dfrac{1}{\sqrt{2}} \\
						\dfrac{1}{\sqrt{2}} &\dfrac{-1}{\sqrt{2}}
					\end{bmatrix} .
					\begin{bmatrix}
						\alpha \\
						\beta
					\end{bmatrix} = 
					\begin{bmatrix}
						\alpha . \dfrac{1}{\sqrt{2}} + \beta.\dfrac{1}{\sqrt{2}} \\
						\alpha . \dfrac{1}{\sqrt{2}} + \beta.\dfrac{-1}{\sqrt{2}}
					\end{bmatrix} = \\
					&\begin{bmatrix}
						\dfrac{1}{\sqrt{2}} . (\alpha + \beta) \\
						\dfrac{1}{\sqrt{2}} . (\alpha - \beta)
					\end{bmatrix}
				\end{aligned}
			\end{equation}

@@

			\begin{equation}
				\label{eq:hadamardApplyedNum}
				\begin{aligned}
					&H|\psi\rangle = \dfrac{1}{\sqrt{2}}.\begin{bmatrix}
						1& 1 \\
						1& -1
					\end{bmatrix} . \begin{bmatrix}
						1 \\
						0
					\end{bmatrix} = \\
					&\begin{bmatrix}
						\dfrac{1}{\sqrt{2}} &\dfrac{1}{\sqrt{2}} \\
						\dfrac{1}{\sqrt{2}} &\dfrac{-1}{\sqrt{2}}
					\end{bmatrix} .
					\begin{bmatrix}
						1 \\
						0
					\end{bmatrix} = 
					\begin{bmatrix}
						1 . \dfrac{1}{\sqrt{2}} + 0.\dfrac{1}{\sqrt{2}} \\
						1 . \dfrac{1}{\sqrt{2}} + 0.\dfrac{-1}{\sqrt{2}}
					\end{bmatrix} = \\
					&\begin{bmatrix}
						\dfrac{1}{\sqrt{2}} . (1 + 0) \\
						\dfrac{1}{\sqrt{2}} . (1 - 0)
					\end{bmatrix} = 
					\begin{bmatrix}
						\dfrac{1}{\sqrt{2}} \\
						\dfrac{1}{\sqrt{2}}
					\end{bmatrix}
				\end{aligned}
			\end{equation}

@@

			\begin{equation}
				\label{eq:rotApplyed}
				Z|\psi\rangle = \begin{bmatrix}
					1& 0 \\
					0& -1
				\end{bmatrix} . \begin{bmatrix}
					\alpha \\
					\beta
				\end{bmatrix} = \begin{bmatrix}
					\alpha \\
					-\beta
				\end{bmatrix}
			\end{equation}

>>>>

			\begin{equation}
				\ket{K} = \vec{K} = \begin{bmatrix}
					a \\
					b \\
					c \\
					d
				\end{bmatrix}
				\label{eq:ketNotationVector}
			\end{equation}

@@

			\begin{equation}
				\ket{\psi} = \alpha . \ket{V_1} + \beta . \ket{V_2}
				\label{eq:ketNotationGeneric}
			\end{equation}

@@

			\par For instance, consider a three-level system. We can define a basis as ${\ket{V_0}, \ket{V_1}, \ket{V_2}}$, where $\ket{V_0}$, $\ket{V_1}$, and $\ket{V_2}$ represent the vectors associated with the respective levels.

@@

			\begin{equation}
				\ket{\psi} = \alpha \ket{V_0} + \beta \ket{V_1} + \gamma \ket{V_2} \qquad,
				\label{eq:multipleLevelState}
			\end{equation}

@@

			\par Being $N=$ the number of levels and $a \in (\alpha, \beta, \gamma, \dots , \omega)$ sometimes the equation \ref{eq:multipleLevelState} can be represented as a sum:

			\begin{equation}
				\ket{\psi} = \sum_{i=1}^{N} a_i . \ket{V_i}= 
				\begin{bmatrix}
					a_{1} \\
					a_{2} \\
					\vdots \\
					a_{n}
				\end{bmatrix}
				\label{eq:multipleLevelStateSum}
			\end{equation}

			\par Some authors prefers to use $\ket{0}, \ket{1}, \ket{2}, \dots , \ket{n} / n < N$ instead of $\ket{V_0},\ket{V_1}, \dots, \ket{V_n} / n < N$.\newline

@@

			\begin{equation}
				\label{eq:linearCombination}
				\begin{aligned}
					\ket{\psi} =
					\begin{bmatrix}
						a_{1} \\
						a_{2} \\
						\vdots \\
						a_{n}
					\end{bmatrix} =
					a_1 . \begin{bmatrix}
						1\\
						0\\
						\vdots \\
						0
					\end{bmatrix}+
					a_2 . \begin{bmatrix}
						0\\
						1\\
						\vdots \\
						0
					\end{bmatrix}+
					\dots + a_n . \begin{bmatrix}
						0\\
						0\\
						\vdots \\
						1
					\end{bmatrix}
				\end{aligned}
			\end{equation}

@@

			\par Briefly, given a ket $\ket{\psi}$ (as shown in equation \ref{eq:ket}), its corresponding bra is denoted by the conjugate transpose of $\ket{\psi}$ (equation \ref{eq:bra}).

			\begin{equation}
				\label{eq:ket}
				\ket{\psi} = 
				\begin{bmatrix}
					u + vo \\
					k + ji \\
					o + li \\
					d + mi
				\end{bmatrix} \in \mathbb{C}
			\end{equation}

			\begin{equation}
				\label{eq:bra}
				\bra{\psi} = [u - vo, k - ji, o - li, d - mi]
			\end{equation}

@@

			\par As we already know, a vector of all probability amplitudes, denoted as $\bra{\psi}$, is given by equation \ref{eq:multipleLevelStateKet}. So, for these "ket" states, there is an associated "bra" notation defined in equation \ref{eq:multipleLevelStateBra} \cite{notacaoDirac}

			\begin{equation}
				\ket{\psi} = \sum_{i=1}^{N} a_i . \ket{V_i} = 
				\begin{bmatrix}
					a_{1} \\
					a_{2} \\
					\vdots \\
					a_{n}
				\end{bmatrix}
				\label{eq:multipleLevelStateKet}
			\end{equation}

			\begin{equation}
				\bra{\psi} = \sum_{j=1}^{N} a_j^* . \bra{V_j} = [a_1^*, a_2^*, \dots, a_n^*]
				\label{eq:multipleLevelStateBra}
			\end{equation}

@@

			\begin{equation}
				\begin{aligned}
					\braket{\psi|\psi} = [a_1^*, a_2^*, \dots, a_n^*] . \begin{bmatrix}
						a_{1} \\
						a_{2} \\
						\vdots \\
						a_{n}
					\end{bmatrix} = &\\ 
					a_1^*.a_1 + a_2^*.a_2 + \dots + a_n^*.a_n = &\\
					P(S_1) + P(S_2) + \dots + P(S_n) = &\\
					\sum_{i=1}^{N} P(S_i)
				\end{aligned}
				\label{eq:multipleLevelStateBraKet}
			\end{equation}

			\par Bra notation allows for the calculation of inner products between vectors. The inner product of a bra vector $\bra{\psi}$ and a ket vector $\ket{\phi}$ is denoted as $\braket{\psi|\phi}$. The inner product yields a complex number, providing information about the similarity or overlap between the two vectors.\newline

			\par For instance, given a vector 
				$\ket{\alpha}= 
				\begin{bmatrix}
					1 \\
					0
				\end{bmatrix}
				$
				 and 
				 $\ket{\beta} = \begin{bmatrix}
					0\\
					1
				\end{bmatrix}$ 
				using the "BraKet" operators to represent the inner product, it will end up like in equation \ref{eq:braket}.

			\begin{equation}
				\label{eq:braket}
				\braket{\alpha|\beta} = 
				[1, 0] .
				\begin{bmatrix}
					0 \\
					1
				\end{bmatrix} = 1.0 + 0.1 = 0
			\end{equation}

@@

		\begin{equation}
			\ket{\psi} = \alpha . \ket{0} + \beta . \ket{1} = 
			\alpha . \begin{bmatrix}
				1 \\
				0
			\end{bmatrix} + \beta .	\begin{bmatrix}
				0 \\
				1
			\end{bmatrix} = 
			\begin{bmatrix}
				\alpha \\
				\beta
			\end{bmatrix}
			\label{eq:qubit}
		\end{equation}

		\par The coefficients $\alpha$ and $\beta$ are complex numbers such that ${|\alpha|}^2 + {|\beta|}^2 = 1$, the vector 
		$\begin{bmatrix}
			\alpha \\
			\beta 
		\end{bmatrix}$ is the qubit itself and, by convention, $\ket{1} = 
		\begin{bmatrix}
			0 \\
			1
		\end{bmatrix} = 1$ and $\ket{0} = 
		\begin{bmatrix}
			1 \\
			0
		\end{bmatrix} = 0$

@@

		\par The Equation \ref{eq:qubit} can be interpreted as follows: The \textbf{non-measured} state of $\psi$ within the vector space defined by vectors $\ket{1}$ and $\ket{0}$ is a linear combination of them  ($\alpha$ times vector $\ket{0}$ plus $\beta$ times vector $\ket{1}$).\newline

@@

			\par As mentioned earlier, quantum gates are represented as matrices, and their number is infinite because the values of the matrix cells can vary across the real numbers ($\mathbb{R}$). Although there is a wide variety of quantum gates and considering $\ket{\psi} = \alpha . \ket{0} + \beta . \ket{1}$, here are a few examples.\newline

@@

			\begin{equation}
				\label{eq:indentityApplyed}
				I\ket{\psi} = \begin{bmatrix}
					1& 0 \\
					0& 1
				\end{bmatrix} . \begin{bmatrix}
					\alpha \\
					\beta
				\end{bmatrix} = \begin{bmatrix}
					\alpha \\
					\beta
				\end{bmatrix} = \ket{\psi}
			\end{equation}

@@

			\begin{equation}
				\label{eq:flipApplyed}
				X\ket{\psi} = \begin{bmatrix}
					0& 1 \\
					1& 0
				\end{bmatrix} . \begin{bmatrix}
					\alpha \\
					\beta
				\end{bmatrix} = \begin{bmatrix}
					\beta \\
					\alpha
				\end{bmatrix}
			\end{equation}

@@

			\begin{equation}
				\label{eq:hadamardApplyed}
				\begin{aligned}
					&H\ket{\psi} = \dfrac{1}{\sqrt{2}}.\begin{bmatrix}
						1& 1 \\
						1& -1
					\end{bmatrix} . \begin{bmatrix}
						\alpha \\
						\beta
					\end{bmatrix} = \\
					&\begin{bmatrix}
						\dfrac{1}{\sqrt{2}} &\dfrac{1}{\sqrt{2}} \\
						\dfrac{1}{\sqrt{2}} &\dfrac{-1}{\sqrt{2}}
					\end{bmatrix} .
					\begin{bmatrix}
						\alpha \\
						\beta
					\end{bmatrix} = 
					\begin{bmatrix}
						\alpha . \dfrac{1}{\sqrt{2}} + \beta.\dfrac{1}{\sqrt{2}} \\
						\alpha . \dfrac{1}{\sqrt{2}} + \beta.\dfrac{-1}{\sqrt{2}}
					\end{bmatrix} = \\
					&\begin{bmatrix}
						\dfrac{1}{\sqrt{2}} . (\alpha + \beta) \\
						\dfrac{1}{\sqrt{2}} . (\alpha - \beta)
					\end{bmatrix}
				\end{aligned}
			\end{equation}

@@

			\begin{equation}
				\label{eq:hadamardApplyedNum}
				\begin{aligned}
					&H\ket{\psi} = \dfrac{1}{\sqrt{2}}.\begin{bmatrix}
						1& 1 \\
						1& -1
					\end{bmatrix} . \begin{bmatrix}
						1 \\
						0
					\end{bmatrix} = \\
					&\begin{bmatrix}
						\dfrac{1}{\sqrt{2}} &\dfrac{1}{\sqrt{2}} \\
						\dfrac{1}{\sqrt{2}} &\dfrac{-1}{\sqrt{2}}
					\end{bmatrix} .
					\begin{bmatrix}
						1 \\
						0
					\end{bmatrix} = 
					\begin{bmatrix}
						1 . \dfrac{1}{\sqrt{2}} + 0.\dfrac{1}{\sqrt{2}} \\
						1 . \dfrac{1}{\sqrt{2}} + 0.\dfrac{-1}{\sqrt{2}}
					\end{bmatrix} = \\
					&\begin{bmatrix}
						\dfrac{1}{\sqrt{2}} . (1 + 0) \\
						\dfrac{1}{\sqrt{2}} . (1 - 0)
					\end{bmatrix} = 
					\begin{bmatrix}
						\dfrac{1}{\sqrt{2}} \\
						\dfrac{1}{\sqrt{2}}
					\end{bmatrix}
				\end{aligned}
			\end{equation}

@@

			\begin{equation}
				\label{eq:rotApplyed}
				Z\ket{\psi} = \begin{bmatrix}
					1& 0 \\
					0& -1
				\end{bmatrix} . \begin{bmatrix}
					\alpha \\
					\beta
				\end{bmatrix} = \begin{bmatrix}
					\alpha \\
					-\beta
				\end{bmatrix}
			\end{equation}
